---
title: 'Building a Geospatial Risk Model: Crime in Chicago'
author: "Allison Carr"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
      include: no
      smooth_scroll: yes
---
##1. Introduction

  Data on demestic violence are some of the most susceptible to selection bias of all crimes. Although domestic violence is, sadly, not an uncommon crime, may instances go unreported. Many victims are unwilling or unable to come forward due to fear and/or social stigma and community norms. In communities where there is mistrust of law enforcement, the reporting situation becomes even more complex. Finally, domestic violence is not a crime that usually occurs on the street, so deterring the crime is not as simple as sending enforcement officers to police a neighborhood. 
  
  This analysis seeks to predict which geographies will be "hotspots" for domestic violence (defined here as domestic battery ONLY - see maps below) based on environmental factors. The goal of using environmental factors to predict crime is to get an idea of where crime may occur even though it is not frequently reported. As the reporting context for domestic violence shows, geospatial risk modeling is a useful approach to consider for preventing crime. Nevertheless, 

###a. Dependent Variable
```{r setup, echo=TRUE, fig.height=8, fig.width=8, fig.align='left', error=FALSE, cache=TRUE, results=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(QuantPsyc)
library(RSocrata)
library(viridis)
library(caret)
library(spatstat)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
library(dplyr)
library(raster)
library(ggplot2)

mapTheme <- function(base_size = 12) {
  theme(
    text = element_text( color = "black"),
    plot.title = element_text(size = 14,colour = "black"),
    plot.subtitle=element_text(face="italic"),
    plot.caption=element_text(hjust=0),
    axis.ticks = element_blank(),
    panel.background = element_blank(),axis.title = element_blank(),
    axis.text = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=2)
  )
}

palette_5_colors <- c("#ff3d6a","#d376a1","#b58fc6","#9aa1e5","#7eb1ff")

policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform(crs=102271) %>%
  dplyr::select(District = dist_num)

policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform(crs=102271) %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))

chicagoBoundary <- 
  st_read("https://data.cityofchicago.org/resource/qqq8-j68g.geojson") %>%
  st_transform(crs = 102271) 

fishnet <- 
  st_make_grid(chicagoBoundary, cellsize = 500) %>%
  st_sf()

fishnet <- 
  fishnet[chicagoBoundary,] %>%
  mutate(uniqueID = rownames(.)) %>%
  dplyr::select(uniqueID)

domestic.battery <- 
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2017/d62x-nvdr") %>% 
  filter(Primary.Type == "BATTERY" & 
           Description == "DOMESTIC BATTERY SIMPLE") %>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Y","X"), sep=",") %>%
  mutate(X = as.numeric(X),
         Y = as.numeric(Y)) %>% 
  na.omit %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant")%>%
  st_transform(102271) %>% 
  distinct()

ggplot() + 
  geom_sf(data = chicagoBoundary) +
  geom_sf(data = domestic.battery, colour="#CC4877", size=0.70, show.legend = "point") +
  labs(title= "Domestic Battery by Point Location", subtitle = "Chicago Open Data, 2017") +
  mapTheme()

crime_net <- 
  domestic.battery %>% 
  dplyr::select() %>% 
  mutate(countBattery = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countBattery = ifelse(is.na(countBattery), 0, countBattery),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = crime_net, aes(fill = countBattery)) +
  scale_fill_viridis(name="", option = "plasma") +
  labs(title = "Domestic Battery Count BY Fishnet") +
  mapTheme()
```

```{r risk factors, echo=TRUE, results=FALSE, warning=FALSE, error=FALSE}
abandonCars <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Abandoned-Vehicles/3c9v-pnva") %>%
  mutate(year = substr(creation_date,1,4)) %>%
  filter(year == "2017") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Abandoned_Cars")

abandonBuildings <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Vacant-and-Abandoned-Building/7nii-7srd") %>%
  mutate(year = substr(date_service_request_was_received,1,4)) %>%
  filter(year == "2017") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Abandoned_Buildings")

graffiti <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Graffiti-Removal-Historical/hec5-y4x5") %>%
  mutate(year = substr(creation_date,1,4)) %>%
  filter(year == "2017") %>%
  filter(where_is_the_graffiti_located_ == "Front" |
           where_is_the_graffiti_located_ == "Rear" | where_is_the_graffiti_located_ == "Side") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Graffiti")

streetLightsOut <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Street-Lights-All-Out/zuxi-7xem") %>%
  mutate(year = substr(creation_date,1,4)) %>%
  filter(year == "2017") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Street_Lights_Out")

sanitation <-
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Sanitation-Code-Complaints-Hi/me59-5fac") %>%
  mutate(year = substr(creation_date,1,4)) %>%
  filter(year == "2017") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Sanitation")

liquorRetail <- 
  read.socrata("https://data.cityofchicago.org/Community-Economic-Development/Business-Licenses-Current-Liquor-and-Public-Places/nrmj-3kcf") %>%
  filter(BUSINESS.ACTIVITY == "Retail Sales of Packaged Liquor") %>%
  dplyr::select(Y = LATITUDE, X = LONGITUDE) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Liquor_Retail")

tobaccoRetail <- 
  read.socrata("https://data.cityofchicago.org/resource/uupf-x98q.json") %>%
  filter(license_description == "Tobacco") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Tobacco_Retail")

pawnRetail <-
  read.socrata("https://data.cityofchicago.org/resource/uupf-x98q.json") %>%
  filter(business_activity_id == "624" | business_activity_id == "732") %>%
  dplyr::select(Y = latitude, X = longitude) %>%
  na.omit() %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform(st_crs(fishnet)) %>%
  mutate(Legend = "Pawn_Retail")

neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson") %>%
  st_transform(st_crs(fishnet)) 

```

###b. Risk Factors

  A number of environmental factors, or "risk factor" values, were engineered using open data from the City of Chicago. The maps below display the values of those risk factor via the fishnet that was created to streamline this analysis. Variables with ".nn" are assessed using nearest neighbor distance, a measure which seeks to avoid the pitfalls of using a pre-defined geography (in this case, fishnet cells) as the unit for defining the data. Nearest neighbor for pawn retail and nearest neighbor for tobacco retail were added to the original list of analysis variables. All of these variables show spatial variation, which is what will help develop a correlation with the domestic battery points shown above.
  
###Risk Factors
  
```{r mapped risk, echo=TRUE, fig.height=10, fig.width=16, fig.align='left', warning=FALSE, results=FALSE}
vars_net <- 
  rbind(abandonCars,streetLightsOut,abandonBuildings,
        liquorRetail, graffiti, sanitation, tobaccoRetail, pawnRetail) %>%
  st_join(., fishnet, join=st_within) %>%
  st_set_geometry(NULL) %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
  full_join(fishnet) %>%
  spread(Legend, count, fill=0) %>%
  st_sf() %>%
  dplyr::select(-`<NA>`) %>%
  na.omit()

nn_function <- function(measureFrom,measureTo,k) {
  measureFrom_Matrix <-
    as.matrix(measureFrom)
  measureTo_Matrix <-
    as.matrix(measureTo)
  nn <-   
    get.knnx(measureTo, measureFrom, k)$nn.dist
  output <-
    as.data.frame(nn) %>%
    rownames_to_column(var = "thisPoint") %>%
    gather(points, point_distance, V1:ncol(.)) %>%
    arrange(as.numeric(thisPoint)) %>%
    group_by(thisPoint) %>%
    summarize(pointDistance = mean(point_distance)) %>%
    arrange(as.numeric(thisPoint)) %>% 
    dplyr::select(-thisPoint) %>%
    pull()
  
  return(output)  
}

vars_net$Abandoned_Buildings.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(abandonBuildings), 3)

vars_net$Abandoned_Cars.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(abandonCars), 3)

vars_net$Graffiti.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(graffiti), 3)

vars_net$Liquor_Retail.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(liquorRetail), 3)

vars_net$Street_Lights_Out.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(streetLightsOut), 3)

vars_net$Sanitation.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(sanitation), 3)

vars_net$Pawn_Retail.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(streetLightsOut), 3)

vars_net$Tobacco_Retail.nn =
  nn_function(st_coordinates(st_centroid(vars_net)), st_coordinates(sanitation), 3)

loopPoint <-
  neighborhoods %>%
  filter(name == "Loop") %>%
  st_centroid()

vars_net$loopDistance =
  st_distance(st_centroid(vars_net),loopPoint) %>%
  as.numeric() 

vars_net.long.new <- 
  vars_net %>%
  dplyr::select(Pawn_Retail.nn, Tobacco_Retail.nn, Liquor_Retail.nn, Abandoned_Cars, Street_Lights_Out, Abandoned_Buildings, Sanitation, Graffiti) %>%
  gather(Variable, value, -geometry, -uniqueID)

vars <- unique(vars_net.long.new$Variable)
mapList <- list()

for(i in vars){
  mapList[[i]] <- 
    ggplot() +
    geom_sf(data = filter(vars_net.long.new, Variable == i), aes(fill=value), colour=NA) +
    scale_fill_viridis(name="", option = "plasma") +
    labs(title=i) +
    mapTheme()}

do.call(grid.arrange,c(mapList, nrow=2))

```

```{r finalNet, echo=TRUE, warning=FALSE, results=FALSE}

final_net <-
  left_join(crime_net, st_set_geometry(vars_net, NULL), by="uniqueID") 

final_net <-
  st_centroid(final_net) %>%
  st_join(., dplyr::select(neighborhoods, name)) %>%
  st_join(., dplyr::select(policeDistricts, District)) %>%
  st_set_geometry(NULL) %>%
  left_join(dplyr::select(final_net, geometry, uniqueID)) %>%
  st_sf() %>%
  na.omit()
```


##2. Spatial Structure

  Local Moran's I is a statistic used to assess the amount of an outcome's clustering or randomness relative to its immediate neighbors. The four maps below show the results of this test in "Local Moran's I" and "P-Value." The P-value was then used to identify hotspots for domestic battery. This is an important metric to consider becuase ideally, the model would have the same predictive power for the hotspots as for all the other areas.

###a. Local Moran's I Statistics, Domestic Battery
```{r morans, echo=TRUE, fig.height=10, fig.width=16, fig.align='left', warning=FALSE}

final_net.nb <- poly2nb(final_net, queen=TRUE)
final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)

final_net.localMorans <- 
  cbind(
    as.data.frame(localmoran(final_net$countBattery, final_net.weights)),
    as.data.frame(final_net, NULL)) %>% 
  st_sf() %>%
  dplyr::select(Battery_Count = countBattery, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z > 0)`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.05, 1, 0)) %>%
  gather(Variable, Value, -geometry)

vars <- unique(final_net.localMorans$Variable)
varList <- list()

for(i in vars){
  varList[[i]] <- 
    ggplot() +
    geom_sf(data = filter(final_net.localMorans, Variable == i), aes(fill = Value), colour=NA) +
    scale_fill_viridis(name="", option = "plasma") +
    labs(title=i) +
    mapTheme()}

do.call(grid.arrange,c(varList, ncol = 4))
```

```{r spatial structure, echo=TRUE, warning=FALSE, results=FALSE}
final_net <-
  final_net %>% 
  mutate(battery.isSig = ifelse(localmoran(final_net$countBattery, 
                                            final_net.weights)[,5] <= 0.0000001, 1, 0)) %>%
  mutate(battery.isSig.dist = nn_function(st_coordinates(st_centroid(final_net)),
                                           st_coordinates(st_centroid(
                                             filter(final_net, battery.isSig == 1))), 1 ))
```

###b. Correlation
  
  These correlation plots show the relationship between the risk factors and number of domestic battery incidents. OVerall, many of these factors have a very similar relationship with domestic battery incidents, so some will be removed due to colinearity.
  
```{r corr, echo=TRUE, fig.align='left', fig.height=16, fig.width=10, warning=FALSE}
correlation.long <-
  st_set_geometry(final_net, NULL) %>%
  dplyr::select(-uniqueID, -cvID, -loopDistance, -name, -District) %>%
  gather(Variable, Value, -countBattery)

correlation.cor <-
  correlation.long %>%
  group_by(Variable) %>%
  summarize(correlation = cor(Value, countBattery, use = "complete.obs"))

ggplot(correlation.long, aes(Value, countBattery)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "#cc4877") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Domestic Battery Count as a Function of Risk Factors")
```

###c. Distribution of Dependent Variable
  
  The distribution of domestic battery incidents by grid cells shows that these incidents do not have a normal distribution. This reinforces our interest in predicting the *number* of incidents, not just the probability that a cell will have an incident. Unfortunately, there will be at least one domestic battery incident in most of the grid cells over the course of a year.
  
```{r hist, echo=TRUE, fig.align='left', fig.height=4, fig.width=4}
ggplot(final_net, aes(countBattery)) + 
  geom_histogram(binwidth = 1) +
  labs(title = "Distribution of Domestic Battery by Grid Cell")
```

##3. Accuracy and Generalizability

  For this model to be useful, it needs to be generalizable across different neighborhood contexts. Two types of cross-validation were used to make sure that this model is generalizable at different scales. One version was "leave one group out" cross validation (LOGO-CV) and the other was a random k-fold. The "MAE by Regression" table below shows that there was only a small difference in error between the two cross-validations, which is a good sign for the model's ability to predict for different geographies.

  
###a. Cross Validation
```{r cross, echo=TRUE, results=FALSE, warning=FALSE}
reg.vars <- c("Abandoned_Buildings.nn", "Abandoned_Cars.nn", "Graffiti.nn", 
              "Liquor_Retail.nn", "Street_Lights_Out.nn", "Sanitation.nn", "Tobacco_Retail.nn", "Pawn_Retail.nn", "loopDistance")

reg.ss.vars <- c("Abandoned_Buildings.nn", "Abandoned_Cars.nn", "Graffiti.nn", "Tobacco_Retail.nn", "Pawn_Retail.nn", "Liquor_Retail.nn", 
                 "Street_Lights_Out.nn", "Sanitation.nn", "loopDistance", 
                 "battery.isSig", "battery.isSig.dist")

crossValidate <- function(dataset, id, dependentVariable, indVariables) {
  
  allPredictions <- data.frame()
  cvID_list <- unique(dataset[[id]])
  
  for (i in cvID_list) {
    
    thisFold <- i
    cat("This hold out fold is", thisFold, "\n")
    
    fold.train <- filter(dataset, dataset[[id]] != thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    fold.test  <- filter(dataset, dataset[[id]] == thisFold) %>% as.data.frame() %>% 
      dplyr::select(id, geometry, indVariables, dependentVariable)
    
    regression <-
      glm(countBattery ~ ., family = "poisson", 
          data = fold.train %>% 
            dplyr::select(-geometry, -id))
    
    thisPrediction <- 
      mutate(fold.test, Prediction = predict(regression, fold.test, type = "response"))
    
    allPredictions <-
      rbind(allPredictions, thisPrediction)
    
  }
  return(st_sf(allPredictions))
}

library(dplyr)
reg.cv <- crossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "countBattery",
  indVariables = reg.vars) %>%
  dplyr::select(cvID = cvID, countBattery, Prediction, geometry)

reg.ss.cv <- crossValidate(
  dataset = final_net,
  id = "cvID",
  dependentVariable = "countBattery",
  indVariables = reg.ss.vars) %>%
  dplyr::select(cvID = cvID, countBattery, Prediction, geometry)

reg.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "countBattery",
  indVariables = reg.vars) %>%
  dplyr::select(cvID = name, countBattery, Prediction, geometry)

reg.ss.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",
  dependentVariable = "countBattery",
  indVariables = reg.ss.vars) %>%
  dplyr::select(cvID = name, countBattery, Prediction, geometry)
```

###b. Impact of Spatial Structure Variables

  The maps "Predicted Domestic Battery by Regression" and table "MAE by Regression" show the difference between the regression with and without the spatial structure variables. Spatial structure helps the model identify the crime "hotspots" the that risk factors cannot pick up alone. Overall, the model with sptial structure predicts +/- 3.7 domestic incidents off of what is observed. A large portion of the cells observe <= 3 domestic battery incidents in a year, so in the context of resource allocation, this is not a very good MAE.
  
```{r accurate, echo=TRUE, fig.height=10, fig.width=16, fig.align='left'}
reg.summary <- 
  rbind(
    mutate(reg.cv,           Error = countBattery - Prediction,
           Regression = "Random k-fold CV: Just Risk Factors"),
    
    mutate(reg.ss.cv,        Error = countBattery - Prediction,
           Regression = "Random k-fold CV: Spatial Structure"),
    
    mutate(reg.spatialCV,    Error = countBattery - Prediction,
           Regression = "Spatial LOGO-CV: Just Risk Factors"),
    
    mutate(reg.ss.spatialCV, Error = countBattery - Prediction,
           Regression = "Spatial LOGO-CV: Spatial Structure")) %>%
  st_sf() 

grid.arrange(
  reg.summary %>%
    ggplot() +
    geom_sf(aes(fill = Prediction)) +
    facet_wrap(~Regression) +
    scale_fill_viridis(option = "plasma") +
    labs(title = "Predicted Domestic Battery by Regression") +
    mapTheme() + theme(legend.position="bottom"),
  
  filter(reg.summary, Regression == "Random k-fold CV: Just Risk Factors") %>%
    ggplot() +
    geom_sf(aes(fill = countBattery)) +
    scale_fill_viridis(option="plasma") +
    labs(title = "Observed battery\n") +
    mapTheme() + theme(legend.position="bottom"), ncol = 2)
```

###c. MAE by Regression
```{r MAE, echo=TRUE, results=TRUE}
st_set_geometry(reg.summary, NULL) %>%
  group_by(Regression) %>% 
  summarize(MAE = round(mean(abs(Prediction - countBattery), na.rm = T),2),
            SD_MAE = round(sd(abs(Prediction - countBattery), na.rm = T),2)) %>% 
  kable(caption = "MAE by Regression") %>%
  kable_styling("striped", full_width = F) %>%
  row_spec(2, color = "white", background = "#cc4877") %>%
  row_spec(4, color = "white", background = "#cc4877") %>%
  row_spec(1, color = "black") %>%
  row_spec(3, color = "black")
```

###d. Neighborhood Context

  Because many majoirty non-white communities in the United States are over-policed, there is typically a noticable racial disparity in crime data. Given this, it is important to assess how a policing model performs across racial groups. The table below shows that in majority non-white neighborhoods, the number of domestic battery incidents is over-estimated, but in majority white-neighborhoods, domenstic battery incidents are under-estimated. This is a troubling finding, because it means that a resource allocation scheme based on this model may not reach the correct communities, or it could reinforce law-enforcement's tendency to over-police the communities of color.
  
```{r CENSUS, echo=TRUE, results=FALSE, warning=FALSE}
tracts17 <- 
  get_acs(geography = "tract", variables = c("B01001_001E","B01001A_001E"), 
          year = 2017, state=17, county=031, geometry=T) %>%
  st_transform(102271)  %>% 
  dplyr::select(variable, estimate, GEOID) %>%
  spread(variable, estimate) %>%
  rename(TotalPop = B01001_001,
         NumberWhites = B01001A_001) %>%
  mutate(percentWhite = NumberWhites / TotalPop,
         raceContext = ifelse(percentWhite > .5, "Majority_White", "Majority_Non_White")) %>%
  .[neighborhoods,]
```

```{r nhood, echo=TRUE, results=FALSE, warning=FALSE}
final_reg <- 
  filter(reg.summary, Regression == "Spatial LOGO-CV: Spatial Structure" |
           Regression == "Spatial LOGO-CV: Just Risk Factors") %>%
  mutate(uniqueID = rownames(.))

final_reg.tracts <- 
  st_join(st_centroid(final_reg), tracts17) %>%
  st_set_geometry(NULL) %>%
  left_join(dplyr::select(final_reg, uniqueID)) %>%
  st_sf() %>%
  na.omit()

st_set_geometry(final_reg.tracts, NULL) %>%
  group_by(Regression, raceContext) %>%
  summarize(mean.Error = mean(Error, na.rm = T)) %>%
  spread(raceContext, mean.Error) %>%
  kable(caption = "Mean Error by Neighborhood Racial Context") %>%
  kable_styling("striped", full_width = F) 
```

##4. Model Performance vs. Buisness as Usual

  One of the most important ways to evaluate a model is to assess whether the allocation scenario presented by the model is better than the "bussiness as usual" allocation scenario. Here we assume that Chicago PD is using a "hotspot" policing model (visualized with kernel density), where more resources are allocated to places where a particular crime is known to occur. With the spatial risk prediction model, the goal is that we can also point resources to the places that have high latent risk, but where the crime is not typically known to occur. Therefore, a good risk prediction model will show more observed incidents of domestic battery.
  
  In the map and plot below, it is apparent that the risk prediction model presents a more nuanced view of grid cells where domestic battery may occur. Per the graph, there is also a higher rate of observed test set crimes in the highest risk category for the predictive model. This means that the model is likely to allocate reosurces better than the business as usual approach indicated by kernel density. 

```{r KERNEL, echo=TRUE, results=FALSE, warning=FALSE}
batt_ppp <- as.ppp(st_coordinates(domestic.battery), W = st_bbox(final_net))
batt_KD.1000 <- spatstat::density.ppp(batt_ppp, 1000)
batt_KD.1500 <- spatstat::density.ppp(batt_ppp, 1500)
batt_KD.2000 <- spatstat::density.ppp(batt_ppp, 2000)
batt_KD.df <- rbind(
  mutate(data.frame(rasterToPoints(mask(raster(batt_KD.1000), as(neighborhoods, 'Spatial')))), Legend = "1000 Ft."),
  mutate(data.frame(rasterToPoints(mask(raster(batt_KD.1500), as(neighborhoods, 'Spatial')))), Legend = "1500 Ft."),
  mutate(data.frame(rasterToPoints(mask(raster(batt_KD.2000), as(neighborhoods, 'Spatial')))), Legend = "2000 Ft.")) 

batt_KD.df$Legend <- factor(batt_KD.df$Legend, levels = c("1000 Ft.", "1500 Ft.", "2000 Ft."))

batt_ppp <- as.ppp(st_coordinates(domestic.battery), W = st_bbox(final_net))
batt_KD <- spatstat::density.ppp(batt_ppp, 1000)

batt_KDE_sf <- as.data.frame(batt_KD) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%

  mutate(label = "Kernel Density",
         Risk_Category = ntile(value, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%")) %>%
  
  bind_cols(
    aggregate(
      dplyr::select(domestic.battery) %>% mutate(battCount = 1), ., length) %>%
      mutate(battCount = replace_na(battCount, 0))) %>%
  
  dplyr::select(label, Risk_Category, battCount)

head(batt_KDE_sf)


batt_risk_sf <-
  filter(final_reg, Regression == "Spatial LOGO-CV: Spatial Structure") %>%
  mutate(label = "Risk Predictions",
         Risk_Category = ntile(Prediction, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%")) %>%
  bind_cols(
    aggregate(
      dplyr::select(domestic.battery) %>% mutate(battCount = 1), ., length) %>%
      mutate(battCount = replace_na(battCount, 0))) %>%
  dplyr::select(label,Risk_Category, battCount)
```

###a. Mapped Comparison
```{r both plot, echo=TRUE, fig.align='left', fig.height=10, fig.width=12, warning=FALSE}
rbind(batt_KDE_sf, batt_risk_sf) %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
  geom_sf(aes(fill = Risk_Category), colour = NA) +
  geom_sf(data = sample_n(domestic.battery, 1500), size = .8, colour = "black") +
  facet_wrap(~label) +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  labs(title="Comparison of Kernel Density and Risk Predictions",
       subtitle="Relative to test set points (in black)") +
  mapTheme()
```

###b. Graphed Comparison
```{r both, echo=TRUE, fig.align='left', fig.height=6, fig.width=8, warning=FALSE}
rbind(batt_KDE_sf, batt_risk_sf) %>%
  st_set_geometry(NULL) %>%
  gather(Variable, Value, -label, -Risk_Category) %>%
  group_by(label, Risk_Category) %>%
  summarize(countBattery = sum(Value)) %>%
  ungroup() %>%
  group_by(label) %>%
  mutate(Rate_of_test_set_crimes = countBattery / sum(countBattery)) %>%
  ggplot(aes(Risk_Category,Rate_of_test_set_crimes)) +
  geom_bar(aes(fill=label), position="dodge", stat="identity") +
  scale_fill_viridis(discrete = TRUE, option = "plasma") +
  labs(title = "Comparison of Kernel Density and Risk Prediction")
```

##5. Conclusion

  There is significant amount of selection bias for domestic violence due to the crime being under-reported for a number of societal factors. In addition to that, different communities are policied differently, complicating the selection bias further. For these reasons, it is useful to attempt to model domestic violence incidents based on environmental risk factors. 

  Interestingly, studies (https://www.jstor.org/stable/10.1525/sp.2004.51.3.326?seq=1#page_scan_tab_contents) have found that rates of domestic violence are the similar or the same across race when you control for envionmental factors. However, because Chicago is such a segregated city, the environments that non-white vs. white people live in are significantly different. This poses a problem when you are predicting based on environmental factors. As the neighborhood context analysis showed, the model's error was different depending on whether a neighborhood was majority white or majority non-white. 
  
  Overall, this model could could be employed as a resource allocation tool for education and community outreach, because it is able to define high-risk areas well, but it should not be employed to dictate where policing occurs. There are many more social factors that should also be considered in addition to the spatial risk factors that have been outlined here.
  